---
title: "Practice Session 8"
format: pdf
editor: source
editor_options: 
  chunk_output_type: console
---

# Intro

The focus of this practice session will be to perform hypothesis tests for the difference of two or more means. We will look at various test statistics that can be used. We will also conduct hypothesis testing for correlations.




## Question : Hypothesis test for one prortion using ** Randominized distribution**, Noraml Distribution **, ** Standard Normal Distribution**


Incentives for Quitting Smoking: Do They Work? Exercise 5.31 describes a study examining incentives to quit smoking. With no incentives, the proportion of smokers trying to quit who are still abstaining six months later is about 0.06. Participants in the study were randomly assigned to one of four different incentives, and the proportion successful was measured six months later. Of the 498 participants in the group with the least success, 47 were still abstaining from smoking six months later. We wish to test to see if this provides evidence that even the smallest incentive works better than the proportion of 0.06 with no incentive at all.

1.) State the null and alternative hypotheses, and give the notation and value of the sample statistic.

2.) Use a randomization distribution and the observed sample statistic to find the p-value.

3.) Give the mean and standard error of the normal distribution that most closely matches the randomization distribution, and then use this normal distribution with the observed sample statistic to find the p-value.

4.) Use the standard error found from the randomization distribution in part (2) to find the standardized test statistic, and then use that test statistic to find the p-value using a standard normal distribution.

5.) Compare the p-values from parts (2), (3), and (4). Use any of these p-values to give the conclusion of the test.

**Answers:**

```{r}
# your code here
```

7.) Calculate p-value 

```{r}
# your code here
```

8.) Make a decision and state your conclusion:

```{r}
#your code here
```

$\\$

## Question 2: Caffeine Taps
```

*b.)* Find some favorites statistics of `Exercise hours` for female and male students. You might find the function: `mosaic::favstats` useful.
*Note*: you can search online for this function arguments.

```{r}
#your code here
```
 
*c.)* Subset the data `ExerciseHours` to two groups: `F` and `M`.


```{r}
#your code here
```

2.) **Step 2**: Compute the observed statistic (mean difference of exercise hours for Female and Male).

```{r}
#your code here
